Motor Learning Without Moving: Proprioceptive and Predictive Hand Localization

### 2019-05-11 paper/paper.Rmd
Plos needed a better ethics statement...

---
title: 'Motor Learning Without Moving: Proprioceptive and Predictive Hand Localization
  After Passive Visuoproprioceptive Discrepancy Training'
author:
- affiliation:
  - YORK
  - MANSOURA
  name: Ahmed A. Mostafa
- affiliation: YORK
  corresponding: thartbm@gmail.com
  email: thartbm@gmail.com
  name: Bernard Marius 't Hart
- affiliation: YORK
  email: deniseh@yorku.ca
  name: Denise Y.P. Henriques
address:
- address: CVR / Kinesiology and Health Science, York University, Toronto, Ontario,
    Canada
  code: YORK
- address: Faculty of Physical Education, Mansoura University, Mansoura, Egypt
  code: MANSOURA
output:
  word_document: default
  rticles::plos_article: default
  pdf_document: default
  html_document:
    df_print: paged
csl: plos.csl
fig_caption: yes
header-includes:
- \usepackage{threeparttable}
- \usepackage[table]{xcolor}
- \usepackage{graphicx}
- \usepackage{tabularx}
- \usepackage{longtable}
- \usepackage{siunitx}
bibliography: bibliography.bib
abstract: |
  An accurate estimate of limb position is necessary for movement planning, before and after motor learning. Where we localize our unseen hand after a reach depends on felt hand position, or proprioception, but in studies and theories on motor adaptation this is quite often neglected in favour of predicted sensory consequences based on efference copies of motor commands. Both sources of information should contribute, so here we set out to further investigate how much of hand localization depends on proprioception and how much on predicted sensory consequences. We use a training paradigm combining robot controlled hand movements with rotated visual feedback that eliminates the possibility to update predicted sensory consequences (‘exposure training’), but still recalibrates proprioception, as well as a classic training paradigm with self-generated movements in another set of participants. After each kind of training we measure participants’ hand location estimates based on both efference-based predictions and afferent proprioceptive signals with self-generated hand movements (‘active localization’) as well as based on proprioception only with robot-generated movements (‘passive localization’). In the exposure training group, we find indistinguishable shifts in passive and active hand localization, but after classic training, active localization shifts more than passive, indicating a contribution from updated predicted sensory consequences. Both changes in open-loop reaches and hand localization are only slightly smaller after exposure training as compared to after classic training, confirming that proprioception plays a large role in estimating limb position and in planning movements, even after adaptation. (data: https://doi.org/10.17605/osf.io/zfdth, preprint: https://doi.org/10.1101/384941)
---

@@

Twenty-five right-handed participants were recruited for the exposure training group in this study. One participant was excluded for not following task instructions, and three were excluded for low performance on a task that ensures attention during the exposure training. All analyses presented here pertain to the remaining twenty-one participants (mean age: 20.1±2.3, age range: 18-25, 13 females), but the data of the three low-performing participants is included in the online dataset @Mostafa2017 ([https://osf.io/zfdth](https://osf.io/zfdth/)). Participants provided prior, written informed consent in accordance with the ethical guidelines set by the York Human subjects Review Subcommittee (#2014-240). All had normal or corrected-to-normal vision, and received credit toward an undergraduate psychology course. Participants were screened verbally, and all reported being right handed and not having any history of visual, neurological, and/or motor dysfunction. Published data from 21 further participants (mean age: 24, age range: 18-38, 11 females) that completed classic visuomotor training is used for comparison @Hart2016.

>>>>

---
title: 'Motor Learning Without Moving: Proprioceptive and Predictive Hand Localization
  After Passive Visuoproprioceptive Discrepancy Training'
author:
- affiliation:
  - YORK
  - MANSOURA
  name: Ahmed A. Mostafa
- affiliation: YORK
  corresponding: thartbm@gmail.com
  email: thartbm@gmail.com
  name: Bernard Marius 't Hart
- affiliation: YORK
  email: deniseh@yorku.ca
  name: Denise Y.P. Henriques
address:
- address: CVR / Kinesiology and Health Science, York University, Toronto, Ontario,
    Canada
  code: YORK
- address: Faculty of Physical Education, Mansoura University, Mansoura, Egypt
  code: MANSOURA
output:
  rticles::plos_article: default
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
csl: plos.csl
fig_caption: yes
header-includes:
- \usepackage{threeparttable}
- \usepackage[table]{xcolor}
- \usepackage{graphicx}
- \usepackage{tabularx}
- \usepackage{longtable}
- \usepackage{siunitx}
bibliography: bibliography.bib
abstract: |
  An accurate estimate of limb position is necessary for movement planning, before and after motor learning. Where we localize our unseen hand after a reach depends on felt hand position, or proprioception, but in studies and theories on motor adaptation this is quite often neglected in favour of predicted sensory consequences based on efference copies of motor commands. Both sources of information should contribute, so here we set out to further investigate how much of hand localization depends on proprioception and how much on predicted sensory consequences. We use a training paradigm combining robot controlled hand movements with rotated visual feedback that eliminates the possibility to update predicted sensory consequences (‘exposure training’), but still recalibrates proprioception, as well as a classic training paradigm with self-generated movements in another set of participants. After each kind of training we measure participants’ hand location estimates based on both efference-based predictions and afferent proprioceptive signals with self-generated hand movements (‘active localization’) as well as based on proprioception only with robot-generated movements (‘passive localization’). In the exposure training group, we find indistinguishable shifts in passive and active hand localization, but after classic training, active localization shifts more than passive, indicating a contribution from updated predicted sensory consequences. Both changes in open-loop reaches and hand localization are only slightly smaller after exposure training as compared to after classic training, confirming that proprioception plays a large role in estimating limb position and in planning movements, even after adaptation. (data: https://doi.org/10.17605/osf.io/zfdth, preprint: https://doi.org/10.1101/384941)
---

@@

Twenty-five right-handed participants were recruited for the exposure training group in this study. One participant was excluded for not following task instructions, and three were excluded for low performance on a task that ensures attention during the exposure training. All analyses presented here pertain to the remaining twenty-one participants (mean age: 20.1±2.3, age range: 18-25, 13 females), but the data of the three low-performing participants is included in the online dataset @Mostafa2017 ([https://osf.io/zfdth](https://osf.io/zfdth/)). Participants provided prior, written informed consent in accordance with the ethical guidelines set by the York Human Participants Review Committee (#2014-240). All had normal or corrected-to-normal vision, and received credit toward an undergraduate psychology course. Participants were screened verbally, and all reported being right handed and not having any history of visual, neurological, and/or motor dysfunction. Published data from 21 further participants (mean age: 24, age range: 18-38, 11 females) that completed classic visuomotor training is used for comparison @Hart2016.
